### 2009-06-16 paper.tex
last edits

A number of leading methods for bioinformatics analysis of structural
RNAs use probabilistic grammars as models for pairs of homologous
RNAs.  We show that any such pairwise grammar can be extended to an
entire phylogeny by treating the pairwise grammar as a machine (a
``transducer'') that models a single ancestor-descendant relationship
in the tree, transforming one RNA structure into another.  In addition
to current applications such as RNA genefinding, homology detection,
alignment and secondary structure prediction, this should permit
probabilistic reconstruction of RNA genes ancestral to many
present-day sequences.  We describe algorithms, software
implementations, and an elementary benchmark of the dependence of the
accuracy of three-taxon RNA reconstruction on the outgroup branch
length.  In the Discussion we consider how the three-taxon RNA
alignment-reconstruction-folding algorithm, which is currently very
computationally-expensive, might be made more efficient so that larger
phylogenies could be considered.

>>>>

A number of leading methods for bioinformatics analysis of structural
RNAs use probabilistic grammars as models for pairs of homologous
RNAs.  We show that any such pairwise grammar can be extended to an
entire phylogeny by treating the pairwise grammar as a machine (a
``transducer'') that models a single ancestor-descendant relationship
in the tree, transforming one RNA structure into another.  In addition
to phylogenetic enhancement of current applications such as RNA genefinding, homology detection,
alignment and secondary structure prediction, this should enable
probabilistic phylogenetic reconstruction of RNA sequences that are ancestral to
present-day genes.  We describe statistical inference algorithms, software
implementations, and a simulation-based comparison
of three-taxon maximum likelihood alignment to several other methods
for aligning three sibling RNAs.
In the Discussion we consider how the three-taxon RNA
alignment-reconstruction-folding algorithm, which is currently very
computationally-expensive, might be made more efficient so that larger
phylogenies could be considered.
